A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa

### 2019-10-16 literature_review.Rmd
general additions and house cleaning

---
title: A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa
author: "Ross Coppin"
output:
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
bibliography: ref_lit_review.bib
---

@@

>>>>

---
title: A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa
author: "Ross Coppin"
output:
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
bibliography: references.bib
---

@@

# Particle dispersion modelling

A study by @lebreton2012 proposed a methododology to track floating debris from source to sink based on discriptions of global waste production and ocean surface currents, using a combination of hydrodynamic and particle dispersion modelling. The particle tracking model applied uses a two stage process, (1) a hydrodynamic model solves the equations of motion to characterise the movement of water within the domain of the model, (2) virtual particles are released into the flow field established in the first step, and particles are allowed to move through hydrodynamic forcing. The authors extracted sea surface currents from the  HYCOM/NCODA ocean circulation modelling system. The HYCOM model is forced by the US Navy's Operational Global Atmospheric Prediction System (NOGAPS). The model also takes into account wind stress, wind speed, heat flux, and precipitation; and was looped five times to accuratley represent ocean circulation patterns for 30 years. The authors used the velocity data aquired from HYCOM and coupled it to the Lagrangian particle tracking model Pol3DD which was used to simulate dispersion.

## Particle dispersion models

The Pol3DD model is able to track virtual particles to simulate water borne dispersion of material. Examples include neutrally buoyant anthropogenic material, larvae, oil spills, outfall discharges and sediment transport [@lebreton2012]. The model uses a second-order accurate advection scheme [@black1990] described as follows:
